---
title: "Modeling Ternary Complex Equilibrium with NIMBLE"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(nleqslv)
library(nimble)
library(basicMCMCplots)
library(coda)
library(plyr)
library(dplyr)
library(ggplot2)
library(GGally)
library(ggmcmc)
library(kableExtra)
library(parallel)
```

# EQUILIBRIUM SYSTEM FUNCTIONS
```{r}
# system of equations describing concentrations at equilibrium 
# A: target protein
# B: degrader
# C: E3 ligase
# K_AB, K_BC: equilibrium dissociation constants of binary complexes
# alpha: cooperativity
equilibrium_f <- function(x, parms) {
  # unpack system parameters
  A_t <- parms[1]
  B_t <- parms[2]
  C_t <- parms[3]
  K_AB <- parms[4]
  K_BC <- parms[5]
  alpha <- parms[6]

  # unknowns are square transformations of the variables 
  A <- x[1]^2
  C <- x[2]^2
  ABC <- x[3]^2

  # analytical expressions relating parameters to unknowns
  F1 <- A + K_BC * ABC / (alpha * C) + ABC - A_t
  F2 <- K_AB * K_BC * ABC / (alpha * A * C) + K_BC * ABC / (alpha * C) + K_AB * ABC / (alpha * A) + ABC - B_t
  F3 <- C + K_AB * ABC / (alpha * A) + ABC - C_t
  return(c(F1 = F1, F2 = F2, F3 = F3))
}

# jacobian of system of equations
# the call signature and arguments are identical to equilibrium_f()
equilibrium_jac <- function(x, parms) {
  # unpack variables
  v1 = x[1]
  v2 = x[2]
  v3 = x[3]
  
  # unknowns are square transformations of the variables
  A = x[1]^2
  C = x[2]^2
  ABC = x[3]^2
  
  # unpack parameters
  K_AB <- parms[4]
  K_BC <- parms[5]
  alpha <- parms[6]

  # D_{ij} contains partial derivative of F[i] with respect to x[j]
  D = rbind( 
    c(2 * v1, -2 * K_BC * ABC / (alpha * v2^3), 2 * K_BC * v3 / (alpha * C) + 2 * v3),
    c(-2 * K_AB * K_BC * ABC / (alpha * v1^3 * C) - 2 * K_AB * ABC / (alpha * v1^3),
      -2 * K_AB * K_BC * ABC / (alpha * A * v2^3) - 2 * K_BC * ABC / (alpha * v2^3),
      2 * K_AB * K_BC * v3 / (alpha * A * C) + 2 * K_BC * v3 / (alpha * C) + 2 * K_AB * v3 / (alpha * A) + 2 * v3),
    c(-2 * K_AB * ABC / (alpha * v1^3), 2 * v2, 2 * K_AB * v3 / (alpha * A) + 2 * v3) )

  return(D)
}

# analytical solution for unknowns when cooperativity equals 1
# result will be used as initial guess for solving equilibrium_f()
noncoop_f <- function(parms) {
  # unpack parameters
  A_t <- parms[1]
  B_t <- parms[2]
  C_t <- parms[3]
  K_AB <- parms[4]
  K_BC <- parms[5]

  A <- A_t - (A_t + B_t + K_AB - sqrt((A_t + B_t + K_AB)**2 - 4 * A_t * B_t)) / 2
  C <- C_t - (C_t + B_t + K_BC - sqrt((C_t + B_t + K_BC)**2 - 4 * C_t * B_t)) / 2
  
  # for B_t >> A_t, expressions for A, C can be negative
  # replace negative values with small positive epsilon 
  # cannot be set to 0 because A, C appear in denominators in equilibrium_f()
  A <- ifelse(A < 0, 1e-19, A)
  C <- ifelse(C < 0, 1e-19, C)

  phi_AB <- A_t - A
  phi_BC <- C_t - C
  ABC <- ifelse(B_t == 0, 0, phi_AB * phi_BC / B_t)
  return(c(A, C, ABC))
}

```

# NIMBLE FUNCTIONS
```{r}
# args: parms = c(A_t, B_t, C_t, K_AB, K_BC, alpha)
# returns: [ABC]
wrap_solve <- function(parms) {
  noncoop_sols <- noncoop_f(head(parms, -1))  # non-cooperative solutions
  init_guess <- sqrt(noncoop_sols)  # square-root transform 
  equilibrium_roots <- nleqslv(
    x = init_guess,
    fn = equilibrium_f,
    jac = equilibrium_jac,
    parms = parms,
    control = list(maxit = 5000)
  )
  coop_sols <- equilibrium_roots$x^2  # square transform to concentration units

  return(coop_sols[3])  # return [ABC]
}

# a nimbleRcall() used in NIMBLE model
nimble_solve <- nimbleRcall(
  prototype = function(parms = double(1)) {},
  returnType = double(0),
  Rfun = "wrap_solve"
)

assign('nimble_solve', nimble_solve, envir = .GlobalEnv)
```

# NANOBRET DATA
```{r}
# Nano-BRET assay data from 4 constructs plus wild type
nanobret_csv <- read.csv("data/corrected_nanobret_df.csv")

sorted_min <- sort(unique(nanobret_csv$Minutes))

nanobret_96 <- nanobret_csv %>% 
  filter(Minutes %in% sorted_min[6:7]) %>%  # filter data from Minute 96 or 97
  group_by(Construct) %>%
  mutate(mBU_corrected = (mBU_corrected - min(mBU_corrected)),  # subtract min mBU within each Construct group
         construct_int = recode(Construct,  # recode constructs as integers
            "VHL_WT SMARCA2_L1415S" = 1,
            "VHL_WT SMARCA2_E1420S" = 2,
            "VHL_R69Q SMARCA2_WT" = 3,
            "VHL_Y112F SMARCA2_WT" = 4,
            "VHL_WT SMARCA2_WT" = 5
            ))  



```

## Plot observed mBU
```{r}
# plot mBU vs. [PROTAC]
ggplot(nanobret_96, aes(x = uM, 
                        y = mBU_corrected, 
                        color = factor(Construct, 
                                       levels = c("VHL_WT SMARCA2_WT",
                                                  "VHL_WT SMARCA2_E1420S",
                                                  "VHL_WT SMARCA2_L1415S",
                                                  "VHL_Y112F SMARCA2_WT",
                                                  "VHL_R69Q SMARCA2_WT"),
                                       ordered = F))) +
  geom_point(size = 5) +
#  geom_line(size = 1) +
  scale_x_log10() +
  guides(color = guide_legend(title = "Constructs")) +
  labs(title = "Nano-BRET assay measurements",
       x = bquote("[ACBI1] (" * mu * "M)"),
       y = "mBU") +
  theme(# text=element_text(size=20), #change font size of all text
        axis.text=element_text(size=18), #change font size of axis text
        axis.title=element_text(size=18), #change font size of axis titles
        plot.title=element_text(size=20), #change font size of plot title
        legend.text=element_text(size=15), #change font size of legend text
        legend.title=element_text(size=18)) #change font size of legend title 
        
ggsave("mBU_VHL_SMARCA2_observed.png",
       path = "plots/",
       width = 12,
       height = 8,
       units = "in")

# set legend labels
ggplot(nanobret_96, aes(x = uM, y = mBU_corrected, color = factor(construct_int, 
                                                                  levels = c(5, 2, 1, 4, 3),
                                                                  ordered = T))) +
  geom_point(size=3) +
  geom_line(size=1) +
  scale_x_log10() +
  guides(color = guide_legend(title = bquote(alpha[i]))) + 
  scale_color_hue(labels = c("WT", "E1420S", "L1415S", "Y112F", "R69Q"))  +
  labs(title = "Nano-BRET assay measurements",
       x = bquote("[ACBI1] (" * mu * "M)"),
       y = "mBU") +
  theme(# text=element_text(size=20), #change font size of all text
        axis.text=element_text(size=18), #change font size of axis text
        axis.title=element_text(size=18), #change font size of axis titles
        plot.title=element_text(size=20), #change font size of plot title
        legend.text=element_text(size=15), #change font size of legend text
        legend.title=element_text(size=18)) #change font size of legend title 

ggsave("mBU_VHL_SMARCA2_observed_by_alpha.png",
       path = "plots/",
       width = 12,
       height = 8,
       units = "in")

```

# NANOBRET NIMBLE MODEL
```{r}
# nimble code
code <- nimbleCode({
  # define priors for parameters
  A_t ~ dexp(1)  # concentrations are defined on R+
  C_t ~ dexp(1)
  constraint ~ dconstraint(0.2*C_t <= A_t & A_t <= 0.5*C_t)  # set a constraint on A_t in terms of C_t
  beta ~ dnorm(0, sd=10)  # beta will be exponentiated, so can be non-positive
  kappa ~ dunif(0, 1)  # kappa is a proportion between 0 and 1
  sigma ~ dexp(10)  # sigma is an error term defined on R+
  for (i in 1:N_constructs) {
    alpha[i] ~ dlnorm(3, sdlog = 0.5)  # informed choice of prior for construct cooperativity
  }
  # set wild type cooperativity if known
  # alpha[N_constructs] <- WT_alpha  

  # calculate likelihood
  for (i in 1:N) {
    B_t[i] <- B_x[i] * kappa  # intracellular concentration of degrader
    c[i] <- construct_int[i]  # index for cooperativity
    ABC[i] <- nimble_solve( c( A_t, B_t[i], C_t, K_AB, K_BC, alpha[c[i]] ) )
    mu[i] <- ABC[i] * exp(beta)  # [ABC] scaled by a factor
    y[i] ~ dnorm(mean = mu[i], sd = sigma)  # normal likelihood
  }
})

# list of constants, units are in uM
# N: number of rows
# N_constructs: number of constructs, excluding wild type
# B_x: vector containing extracellular concentrations of degrader
# construct_int: vector containing integer indices i corresponding to alpha[i]
# K_AB, K_BC: equilibrium dissociation constants of binary complexes
# WT_alpha: cooperativity of wild type proteins
constants <- list(
  N = nrow(nanobret_96),
  N_constructs = length(unique(nanobret_96$construct_int)),
  B_x = nanobret_96$uM,
  construct_int = nanobret_96$construct_int,
  K_AB = 9.26,  # PROTAC to target
  K_BC = 0.0694  # PROTAC to E3
  # WT_alpha = 26
)

# list of data
# y: vector of ternary complex concentrations scaled by factor
data <- list(
  y = nanobret_96$mBU_corrected,
  constraint = 1  # constraint on A_t is set to true
)

# list generates initial values
inits_f <- function() {
  C_t <- runif(n = 1, min = 0, max = 5)  # draw C_t
  
  list(
    C_t = C_t,
    A_t = 0.35*C_t,  # define A_t in terms of C_t
    beta = rnorm(n = 1, mean = 0, sd = 1),
    kappa = runif(n = 1, min = 0.5, max = 1),
    alpha = rnorm(n = 5, mean = 20, sd = 5),  # vector of construct cooperativities
    sigma = rexp(n = 1, rate = 10)
  )
}
```

# MAKE CLUSTER FOR PARALLEL
```{r}
this_cluster <- makeCluster(4)
```

# PARALLEL NIMBLE WRAPPER
```{r}
set.seed(0)

run_MCMC_all_code <- function(seed, code, constants, data, inits_f) {
  library(nimble)
  library(coda)
  library(nleqslv)
  
  equilibrium_f <- function(x, parms) {
    # unpack system parameters
    A_t <- parms[1]
    B_t <- parms[2]
    C_t <- parms[3]
    K_AB <- parms[4]
    K_BC <- parms[5]
    alpha <- parms[6]
  
    # unknowns are square transformations of the variables 
    A <- x[1]^2
    C <- x[2]^2
    ABC <- x[3]^2
  
    # analytical expressions relating parameters to unknowns
    F1 <- A + K_BC * ABC / (alpha * C) + ABC - A_t
    F2 <- K_AB * K_BC * ABC / (alpha * A * C) + K_BC * ABC / (alpha * C) + K_AB * ABC / (alpha * A) + ABC - B_t
    F3 <- C + K_AB * ABC / (alpha * A) + ABC - C_t
    return(c(F1 = F1, F2 = F2, F3 = F3))
  }
  assign('equilibrium_f', equilibrium_f, envir = .GlobalEnv)
  
  equilibrium_jac <- function(x, parms) {
    # unpack variables
    v1 = x[1]
    v2 = x[2]
    v3 = x[3]
    
    # unknowns are square transformations of the variables
    A = x[1]^2
    C = x[2]^2
    ABC = x[3]^2
    
    # unpack parameters
    K_AB <- parms[4]
    K_BC <- parms[5]
    alpha <- parms[6]
  
    # D_{ij} contains partial derivative of F[i] with respect to x[j]
    D = rbind( 
      c(2 * v1, -2 * K_BC * ABC / (alpha * v2^3), 2 * K_BC * v3 / (alpha * C) + 2 * v3),
      c(-2 * K_AB * K_BC * ABC / (alpha * v1^3 * C) - 2 * K_AB * ABC / (alpha * v1^3),
        -2 * K_AB * K_BC * ABC / (alpha * A * v2^3) - 2 * K_BC * ABC / (alpha * v2^3),
        2 * K_AB * K_BC * v3 / (alpha * A * C) + 2 * K_BC * v3 / (alpha * C) + 2 * K_AB * v3 / (alpha * A) + 2 * v3),
      c(-2 * K_AB * ABC / (alpha * v1^3), 2 * v2, 2 * K_AB * v3 / (alpha * A) + 2 * v3) )
  
    return(D)
  }
  assign('equilibrium_jac', equilibrium_jac, envir = .GlobalEnv)

  noncoop_f <- function(parms) {
    # unpack parameters
    A_t <- parms[1]
    B_t <- parms[2]
    C_t <- parms[3]
    K_AB <- parms[4]
    K_BC <- parms[5]
  
    A <- A_t - (A_t + B_t + K_AB - sqrt((A_t + B_t + K_AB)**2 - 4 * A_t * B_t)) / 2
    C <- C_t - (C_t + B_t + K_BC - sqrt((C_t + B_t + K_BC)**2 - 4 * C_t * B_t)) / 2
    
    # for B_t >> A_t, expressions for A, C can be negative
    # replace negative values with small positive epsilon 
    # cannot be set to 0 because A, C appear in denominators in equilibrium_f()
    A <- ifelse(A < 0, 1e-19, A)
    C <- ifelse(C < 0, 1e-19, C)
  
    phi_AB <- A_t - A
    phi_BC <- C_t - C
    ABC <- ifelse(B_t == 0, 0, phi_AB * phi_BC / B_t)
    return(c(A, C, ABC))
  }
  assign('noncoop_f', noncoop_f, envir = .GlobalEnv)

  wrap_solve <- function(parms) {
    noncoop_sols <- noncoop_f(head(parms, -1))  # non-cooperative solutions
    init_guess <- sqrt(noncoop_sols)  # square-root transform 
    equilibrium_roots <- nleqslv(
      x = init_guess,
      fn = equilibrium_f,
      jac = equilibrium_jac,
      parms = parms,
      control = list(maxit = 5000)
    )
    coop_sols <- equilibrium_roots$x^2  # square transform to concentration units
  
    return(coop_sols[3])  # return [ABC]
  }
  assign('wrap_solve', wrap_solve, envir = .GlobalEnv)
  
  # a nimbleRcall() used in NIMBLE model
  nimble_solve <- nimbleRcall(
    prototype = function(parms = double(1)) {},
    returnType = double(0),
    Rfun = "wrap_solve"
  )
  assign('nimble_solve', nimble_solve, envir = .GlobalEnv)
  
  # build the R model object
  Rmodel <- nimbleModel(
    code = code,
    constants = constants,
    data = data,
    inits = inits_f()
  )
  
  conf <- configureMCMC(Rmodel)

  Rmcmc <- buildMCMC(conf)

  Cmodel <- compileNimble(Rmodel)
  Cmcmc <- compileNimble(Rmcmc, project = Rmodel)

  results <- runMCMC(Cmcmc, niter = 10000, nburnin = 1000, thin = 2, 
                     samplesAsCodaMCMC = TRUE, setSeed = seed)
}
```

# RUN CHAINS IN PARALLEL
```{r}
chain_output <- parLapply(cl = this_cluster, 
                          X = 1:4, 
                          fun = run_MCMC_all_code, 
                          code = code,
                          constants = constants,
                          data = data,
                          inits_f = inits_f)

stopCluster(this_cluster)
```

# SEQUENTIAL CHAINS
```{r}
# build the R model object
Rmodel <- nimbleModel(
  code = code,
  constants = constants,
  data = data,
  inits = inits_f()
)

# sum of log probabilities from each node in model
# should be real and negative
Rmodel$calculate()

```

# MCMC configuration and building
```{r}
conf <- configureMCMC(Rmodel)

Rmcmc <- buildMCMC(conf)
```

# compile to C++, and run
```{r}
Cmodel <- compileNimble(Rmodel)
Cmcmc <- compileNimble(Rmcmc, project = Rmodel)
```

# run multiple MCMC chains
```{r}
samples_n4 <- runMCMC(Cmcmc, nchains = 4, 
                      niter = 10000, nburnin = 1000, thin = 2, 
                      inits = inits_f, samplesAsCodaMCMC = TRUE)
```

# SAVING AND READING MCMC RESULTS
```{r}
# saving and reading sampling results to MCMC folder 

# saveRDS(samples_n4, file = "MCMC/nimble_fit_full.rds")

samples_aFix <- readRDS(file = "MCMC/nimble_fit_full_fix_WT_alpha.rds")
samples_aFree <- readRDS(file = "MCMC/nimble_fit_full.rds")

chain_output <- samples_aFix
```

# Summarize Parameter Posterior Samples 
```{r}
names(chain_output) <- c("chain1", "chain2", "chain3", "chain4")

# combine samples from all chains
samples_all <- do.call(rbind, chain_output)
# samples_all <- do.call(rbind, samples_n4)  

ss_all <- samplesSummary(samples_all)  # summary statistics of combined samples

tab <- ss_all[ , c("Mean", "Median", "95%CI_low", "95%CI_upp")]

tab_t <- data.frame(t(tab))  # Transpose for easy iteration
tab_t <- mapply(  # Iterate over each column
  function(df, digits) {
    formatC(df, format="f", digits=digits)
  },
  df=tab_t, digits=c(3, 3, rep(0, 5), 2, 2, 2)
)

tab_t2 <- data.frame(t(tab_t))  # Transpose back
# Re-create names
rownames(tab_t2) <- c("[A]$_t$", "[C]$_t$", 
                      "$\\alpha_{L1415S}$", "$\\alpha_{E1420S}$", "$\\alpha_{R69Q}$", "$\\alpha_{Y112F}$", "$\\alpha_{WT}$",
                      "$\\beta$", "$\\kappa$", "$\\sigma$"
                   )
colnames(tab_t2) <- colnames(tab)


kable(tab_t2, caption = "Mean, Median and 95% Bayesian Credible Intervals of Model Parameters") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, position = "center")
```

## Effective Size in chains
```{r}
lapply(chain_output, effectiveSize)
```

```{r}
# samplesPlot(samples_n4, var = "alpha")
# samplesPlot(samples_n4, var = c('A_t', 'C_t'))
# samplesPlot(samples_n4, var = c('kappa', 'sigma'))
# samplesPlot(samples_n4, var = 'beta')
```

## Gelman-Rubin diagnostic
```{r}
# Gelman-Rubin diagnostics assess parameter sampling convergence between chains
GR.diag <- gelman.diag(chain_output, multivariate = FALSE)
GR.diag
```

## GGS object
```{r}
# ggs.nanobret <- ggs(samples_n4)
ggs.nanobret <- ggs(chain_output)

ggs.nanobret$Parameter <- plyr::revalue(ggs.nanobret$Parameter, 
                                        c("alpha[1]" = "alpha[L1415S]",
                                          "alpha[2]" = "alpha[E1420S]",
                                          "alpha[3]" = "alpha[R69Q]",
                                          "alpha[4]" = "alpha[Y112F]",
                                          "alpha[5]" = "alpha[WT]"))

```


## alpha plots
```{r}
ggs.nanobret %>% 
  # filter(Parameter %in% c("alpha[L1415S]", "alpha[E1420S]", "alpha[R69Q]", "alpha[Y112F]")) %>%
  ggs_traceplot(family = "alpha", greek = TRUE) +
  theme_bw()

ggs.nanobret %>%
  # filter(Parameter %in% c("alpha[L1415S]", "alpha[E1420S]", "alpha[R69Q]", "alpha[Y112F]")) %>%
  ggs_density(family = "alpha", greek = T) +
  theme_bw()
```

## kappa, sigma plots
```{r}
ggs.nanobret %>% 
  filter(Parameter %in% c("kappa", "sigma")) %>%
  ggs_traceplot(greek = TRUE) +
  theme_bw()

ggs.nanobret %>%
  filter(Parameter %in% c("kappa", "sigma")) %>%
  ggs_density(greek = T) +
  theme_bw()
```

## A_t, C_t, beta plots
```{r}
ggs.nanobret %>% 
  filter(Parameter %in% c("A_t", "C_t", "beta")) %>%
  mutate(Parameter = recode_factor(Parameter, A_t = "A[t]", C_t = "C[t]")) %>%
  ggs_traceplot(greek = TRUE, ) +
  theme_bw()

ggs.nanobret %>% 
  filter(Parameter %in% c("A_t", "C_t", "beta")) %>%
  mutate(Parameter = recode_factor(Parameter, A_t = "A[t]", C_t = "C[t]")) %>%
  ggs_density(greek = TRUE, ) +
  theme_bw()
```

## alpha running mean plots
```{r}
ggs.nanobret %>%
  # filter(Parameter %in% c("alpha[L1415S]", "alpha[E1420S]", "alpha[R69Q]", "alpha[Y112F]")) %>%
  ggs_running(family = "alpha", greek = TRUE) + 
  theme_bw()
```

## GGS objects from fixed and free WT alpha models
```{r}
ggs.fix <- ggs(samples_aFix, 
               family = "alpha"
               )
ggs.free <- ggs(samples_aFree, 
                family = "alpha"
                )

ggs.fix$Parameter <- plyr::revalue(ggs.fix$Parameter, 
                                    c("alpha[1]" = "alpha[L1415S]",
                                      "alpha[2]" = "alpha[E1420S]",
                                      "alpha[3]" = "alpha[R69Q]",
                                      "alpha[4]" = "alpha[Y112F]",
                                      "alpha[5]" = "alpha[WT]"
                                      )
                                   )
ggs.free$Parameter <- plyr::revalue(ggs.free$Parameter, 
                                    c("alpha[1]" = "alpha[L1415S]",
                                      "alpha[2]" = "alpha[E1420S]",
                                      "alpha[3]" = "alpha[R69Q]",
                                      "alpha[4]" = "alpha[Y112F]",
                                      "alpha[5]" = "alpha[WT]"
                                      )
                                   )
```


## alpha caterpillar plots


```{r}
ggs.fix %>%
  filter(Parameter != "alpha[WT]") %>%
  ggs_caterpillar(greek = T) +
  theme_bw() +
  labs(x="", y="Mutant cooperativity") +
  geom_vline(xintercept = 26, size = 1, color = "tomato") +
  geom_text(aes(x = 26, label = "alpha[WT]", y = 1), 
            parse = T, colour="tomato", size = 6, nudge_x = 2.3) +
  aes(color = factor(Parameter, 
                     levels = c("alpha[WT]",
                                "alpha[E1420S]",
                                "alpha[L1415S]",
                                "alpha[Y112F]",
                                "alpha[R69Q]"),
                     ordered = F)) +
  guides(color = guide_legend(title = "Cooperativity")) +
  theme(# text=element_text(size=20), #change font size of all text
      axis.text=element_text(size=13), #change font size of axis text
      axis.title=element_text(size=13), #change font size of axis titles
      legend.position="none", #change font size of legend text
      ) 

ggsave("caterpillar_plot_fixed_WT_alpha.png", path = "plots/", width = 4, height = 3.5, units = "in")
```

```{r}
ggs.free %>%
  ggs_caterpillar(greek = T) +
  theme_bw() +
  labs(title = "Highest Posterior Density Intervals",
       subtitle = "Thick lines: 90%, Thin lines: 95%") +
  geom_vline(xintercept = 26, size = 1, color = "tomato") +
  geom_text(aes(x = 26, label = "alpha[WT]", y = 1), 
            parse = T, colour="tomato", size = 5, nudge_x = 1.5) +
  aes(color = factor(Parameter, 
                     levels = c("alpha[WT]",
                                "alpha[E1420S]",
                                "alpha[L1415S]",
                                "alpha[Y112F]",
                                "alpha[R69Q]"),
                     ordered = F)) +
  guides(color = guide_legend(title = "Cooperativity")) +
  theme(# text=element_text(size=20), #change font size of all text
      axis.text=element_text(size=18), #change font size of axis text
      axis.title=element_text(size=18), #change font size of axis titles
      plot.title=element_text(size=20), #change font size of plot title
      plot.subtitle=element_text(size=15),
      legend.text=element_text(size=15), #change font size of legend text
      legend.title=element_text(size=18)) #change font size of legend title 
ggsave("caterpillar_plot_free_WT_alpha.png", path = "plots/", width = 12, height = 8, units = "in")

ggs_caterpillar(D = list(ggs.fix, ggs.free),
                model_labels = c("fixed alpha[WT]", "free alpha[WT]"),
                greek = F) + 
  theme_bw() +
  labs(title = "Highest Posterior Density Intervals",
       subtitle = "Thick lines: 90%, Thin lines: 95%") +
  geom_vline(xintercept = 26, size = 1, color = "tomato") +
  geom_text(aes(x = 26, label = "alpha[WT]", y = 1), 
            parse = T, colour="tomato", size = 5, nudge_x = 2.5) +
  aes(color = factor(Parameter, 
                     levels = c("alpha[WT]",
                                "alpha[E1420S]",
                                "alpha[L1415S]",
                                "alpha[Y112F]",
                                "alpha[R69Q]"),
                     ordered = F)) +
  guides(color = guide_legend(title = "Cooperativity")) +
  theme(# text=element_text(size=20), #change font size of all text
      axis.text=element_text(size=15), #change font size of axis text
      axis.title=element_text(size=18), #change font size of axis titles
      plot.title=element_text(size=20), #change font size of plot title
      plot.subtitle=element_text(size=15),
      legend.text=element_text(size=15), #change font size of legend text
      legend.title=element_text(size=18)) #change font size of legend title 
ggsave("caterpillar_plot_fixed_free_WT_alpha.png", path = "plots/", width = 12, height = 8, units = "in")

```

## A_t, C_t, beta correlation plots
```{r}
labels <- c(A_t = "A[t]", C_t = "C[t]", beta = "beta")

ggs.nanobret %>%
  filter(Parameter %in% c("A_t", "C_t", "beta")) %>%
  mutate(Parameter = recode_factor(Parameter, A_t = "A[t]", C_t = "C[t]")) %>%
  ggs_pairs(mapping = aes(color = Chain),
            lower = list(continuous = wrap("points", alpha = 0.1)),
            diag = list(continuous = wrap("densityDiag", alpha = 0.5, color = NA)),
            labeller = "label_parsed") +
  theme_bw() 
```

# TESTING ROOT-FINDING 
```{r}
# prints non-cooperative solutions and results of cooperative system root-finding
test_roots <- function(parms) {
  # parms = c(A_t, B_t, C_t, K_AB, K_BC, alpha)
  noncoop_sols <- noncoop_f(head(parms, -1))
  init_guess <- sqrt(noncoop_sols)  # square-root transform non-cooperative solutions
  equilibrium_roots <- nleqslv(
    x = init_guess,
    fn = equilibrium_f,
    jac = equilibrium_jac,
    parms = parms,
    control = list(maxit = 5000)
  )
  coop_sols <- equilibrium_roots$x^2  # square transform roots to concentration units
  print(noncoop_sols)
  print(coop_sols)
  print(equilibrium_roots$fvec)  # the value of system F at roots
}

# UNIT TEST
# set alpha to 1
# verify with Douglass et al. Supplementary Excel 
test_roots(c(5e-11, 0, 1e-5, 5e-8, 1e-9, 1))
test_roots(c(5e-11, 7.5e-2, 1e-5, 5e-8, 1e-9, 1))
test_roots(c(5e-11, 1.2e-2, 1e-5, 5e-8, 1e-9, 1))
test_roots(c(5e-11, 9.18e-9, 1e-5, 5e-8, 1e-9, 1))
test_roots(c(5e-11, 6.24e-15, 1e-5, 5e-8, 1e-9, 1))
test_roots(c(5e-11, 5, 1e-5, 5e-8, 1e-9, 1))
test_roots(c(5e-11, 1.12e-6, 1e-5, 1.8e-6, 2.5e-7, 1))

# as alpha gets larger, ABC increases and reduces A more than C
test_roots(c(5e-11, 1.12e-6, 1e-5, 1.8e-6, 2.5e-7, 30))
test_roots(c(5e-11, 1.12e-6, 1e-5, 1.8e-6, 2.5e-7, 50))
```

# Testing NIMBLE function
```{r}
# test the nimbleRcall() in a NIMBLE model
testModelCode <- nimbleCode({
    ABC[1] <- nimble_solve(x[1:6])
})

testModel <- nimbleModel(testModelCode, check = FALSE, calculate = FALSE)

ctestModel <- compileNimble(testModel)

## set x:
ctestModel$x <- c(5e-11, 2.81, 1e-5, 1.8e-6, 2.5e-7, 50)

## Answer directly:
wrap_solve(ctestModel$x)

## Answer via the nimble model:
ctestModel$calculate('ABC[1]')
ctestModel$ABC  # This result should match the answer calculated directly
```

# Predict mBU
```{r}
A_t <- 0.013
C_t <- 0.034
beta <- 8.1  # mean 8.58
kappa <- 0.95
alpha <- c(17, 27, 13, 15, 26)
K_AB <- 9.26  # PROTAC to target
K_BC <- 0.0694  # PROTAC to E3

B_tArray <- 10 ^ seq(log10(0.0001), log10(10), length.out = 100)
constructArray <- 1:5
constructNames <- c(
  "VHL_WT SMARCA2_L1415S", 
  "VHL_WT SMARCA2_E1420S", 
  "VHL_R69Q SMARCA2_WT", 
  "VHL_Y112F SMARCA2_WT", 
  "VHL_WT SMARCA2_WT"
  )

N_B_t <- length(B_tArray)
N_construct <- length(constructArray)

nanobret_96_fit <- data.frame(
  uM = numeric(N_B_t * N_construct), 
  mBU_predicted = numeric(N_B_t * N_construct), 
  Construct = character(N_B_t * N_construct)
)

for (i in 1:N_construct) {
  for (j in 1:N_B_t) {
    B_t <- B_tArray[j]
    constructInt <- constructArray[i]
    
    parms <- c(A_t, B_t, C_t, K_AB, K_BC, alpha[constructInt])
    ABC <- wrap_solve(parms)
    mBU <- ABC * exp(beta)

    nanobret_96_fit[(i-1)*N_B_t + j, ] <- c(B_t, mBU, constructNames[constructInt])
  }
}

nanobret_96_fit$uM <- as.numeric(nanobret_96_fit$uM)
nanobret_96_fit$mBU_predicted <- as.numeric(nanobret_96_fit$mBU_predicted)
```

## Combine observed and predicted mBU
```{r}
nanobret_96_obs <- nanobret_96 %>%
  select(uM, mBU_corrected, Construct)

nanobret_96_combined <- merge.data.frame(nanobret_96_obs, nanobret_96_fit, by = "Construct")
head(nanobret_96_combined)

write.csv(nanobret_96_combined, file="data/fitted_nanobret.csv", row.names = FALSE)
```

## Plot fitted and observed mBU
```{r}
library(scales)
ggplot(data = nanobret_96_combined, aes(x = uM.y, y = mBU_predicted, color = factor(Construct, 
    levels = c(
      "VHL_WT SMARCA2_WT",
      "VHL_WT SMARCA2_E1420S",
      "VHL_WT SMARCA2_L1415S",
      "VHL_Y112F SMARCA2_WT",
      "VHL_R69Q SMARCA2_WT"
    ), 
    ordered = F
  ))) +
  theme_bw() +
  geom_line(size = 1, aes(alpha = "fitted")) + 
  geom_point(size = 2, aes(x = uM.x, y = mBU_corrected, alpha = "observed")) +
  scale_x_log10(limits=c(1e-4, 10), breaks=c(1e-4, 1e-3, 1e-2, 0.1, 1, 10), labels=trans_format('log10',math_format(10^.x))) +
  guides(color = guide_legend(title = "Mutant")) +
  scale_color_hue(labels = c("WT", "E1420S", "L1415S", "Y112F", "R69Q")) +
  scale_alpha_manual(name = NULL, values = c(1, 1),
                     guide = guide_legend(override.aes = list(linetype = c(1, 0), shape = c(NA, 16) ))) +
  labs(x = bquote("ACBI1 (" * mu * "M)"), y = "mBU") +
  theme(# text=element_text(size=20), #change font size of all text
        axis.text=element_text(size=13), #change font size of axis text
        axis.title=element_text(size=13), #change font size of axis titles
        legend.text=element_text(size=9), #change font size of legend text
        legend.title=element_text(size=9),
        legend.position = c(.02, .98),
        legend.justification = c("left", "top"),
        legend.box.just = "left",
        # legend.background = element_rect(fill='transparent'), #transparent legend bg
        # legend.box.background = element_rect(fill='transparent'), #transparent legend panel,
        ) #change font size of legend title

ggsave("mBU_VHL_SMARCA2_fixed_WT_alpha_fit.png", path = "plots/", width = 4, height = 3.5, units = "in")
```


